\documentclass{article}

%opening
%\title{}
%\author{}

\begin{document}
Some text here common to both documents

Every year in the United States, natural hazard events threaten lives and livelihoods, resulting in deaths and billions of dollars in damage. The Firm works with many partners to monitor, assess, and conduct targeted research on a wide range of natural hazards so that policymakers and the public have the understanding they need to enhance preparedness, response and resilience.


\textbf{GUE Tech}
\begin{description}
\item[\underline{\mdseries 1 PhD student position (NN) TV-L E13 (75\%) for 36 months}:] 
The PhD will carry out the tasks in WP2 involving the cross-correlation and array-based relocations;  spatio-temporal analysis search for fluid signatures and stress; time-dependent Vp/Vs ratio determination  
\item[\underline{\mdseries 1 student research assistant (40 hours/month for 12 months)}:] Support in processing of seismic data: event detection/selection, phase picking, quality control and data archiving.
\end{description}
%\mbox{}\hfill$3\times 6000.00/a = 18000.00$\texteuro
\textbf{Starfleet Academy} %\hspace*{1cm}\\[-\baselineskip]
\begin{description}\item[\underline{\mdseries 1 postdoc position (M. Mouse) TV-L E13 for 24 months}:] 
The postdoc will carry out tasks in WP1 involving the calculation of ambient noise auto- and cross-correlation functions and their analysis in terms of static and time-variable velocity; analysis of multiplet coda waves in terms of velocity changes; inference of source regions and physical mechanisms of velocity changes. M. Mouse is a highly experienced seismologist with an excellent knowledge of the study area who coordinated (alongside D. Duck) and contributed substantially to the field acquisition in this project.  
%: Spatio-temporal distribution of source parameters from earthquake data\\
%\mbox{}\hfill$3\times 48814.60/a = 146443.80$\texteuro\\[0.3cm]
\item[\underline{\mdseries 1 student research assistant (40 hours/month for 12 months)}:] Support in processing of seismic data, preparation of presentation material/posters.
%\mbox{}\hfill$3\times 6000.00/a = 18000.00$\texteuro
\end{description}
%

A list with a removed item.
\begin{itemize}
\item The United States Geological Survey (USGS, formerly simply Geological Survey) is a scientific agency of the United States Government. 
\end{itemize}
and at the beginning
\begin{itemize}
\item The USGS is a fact-finding research organization with no regulatory responsibility.
\end{itemize}

And a few word to finish off in a consistent manner.
\end{document}

