
\documentclass{article}
%\usepackage[
%        svgnames,
%            ]{xcolor}
%\usepackage[
%    color=DodgerBlue,
%        ]{siunitx}
\usepackage{siunitx}

\begin{document}

Aaa \ang{6;7;6.5} bbb ccc ddd eee fff ggg hhh iii jjj kkk lll mmm nnn ooo ppp
qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa \num{845} bbb ccc ddd eee fff ggg
hhh iii jjj kkk lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz.
Aaa \si{\joule\per\kelvin} bbb ccc ddd eee fff ggg hhh iii jjj kkk lll mmm nnn
ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa \SI{4}{\meter\per\second}
bbb ccc ddd eee fff ggg hhh iii jjj kkk lll mmm nnn ooo ppp qqq rrr sss ttt uuu
vvv www xxx yyy zzz. Aaa \numlist{1;2;3} bbb ccc ddd eee fff ggg hhh iii jjj
kkk lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\numrange{10}{20} bbb ccc ddd eee fff ggg hhh iii jjj kkk lll mmm nnn ooo ppp
qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\SIlist{1;2;3}{\watt\per\square\meter\per\kelvin} bbb ccc ddd eee fff ggg hhh
iii jjj kkk lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\SIrange{5}{12}{\bar} bbb ccc ddd eee fff ggg hhh iii jjj kkk lll mmm nnn ooo
ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz.

Aaa \ang[arc-separator = \,]{6;7;6.5} bbb ccc ddd eee fff ggg hhh iii jjj kkk
lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\num[round-mode=figures]{845} bbb ccc ddd eee fff ggg hhh iii jjj kkk lll mmm
nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\si[per-mode=symbol]{\joule\per\kelvin} bbb ccc ddd eee fff ggg hhh iii jjj kkk
lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\SI{4}{\meter\per\second} bbb ccc ddd eee fff ggg hhh iii jjj kkk
lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\numlist[list-separator = {; }]{1;2;3} bbb ccc ddd eee fff ggg hhh iii jjj kkk
lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\numrange[range-phrase = --]{10}{20} bbb ccc ddd eee fff ggg hhh iii jjj kkk
lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx yyy zzz. Aaa
\SIlist[list-units = brackets]{1;2;3}{\watt\per\square\meter\per\kelvin} bbb
ccc ddd eee fff ggg hhh iii jjj kkk lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv
www xxx yyy zzz. Aaa \SIrange[range-units = brackets]{5}{12}{\bar} bbb ccc ddd
eee fff ggg hhh iii jjj kkk lll mmm nnn ooo ppp qqq rrr sss ttt uuu vvv www xxx
yyy zzz.

\end{document}
